Learning Auditory Similarities Between Instrumental Playing Techniques

### 2019-10-31 paper/paper.tex
fix sections

In summary, audio signals from isolated musical notes may vary across three categorical variables (instrument, mute, and technique) and two continuous variables (intensity and pitch).
The Cartesian product of all five variables results in a combinatorial explosion of possible timbral sensations.
The Studio On Line dataset (SOL), recorded at Ircam in 2000, offers a dense sampling of this combinatorial explosion, amounting to X audio signals in total. % TODO check year in which SOL was recorded, update with number
We should note that this dense sampling comes at the detriment of other factors of acoustic variability, such as identity of performer, identity of instrument manufacturer, audio acquisition equipment, and room response characteristics; which are all restricted to singletons.
Addressing these factors of variability is besides the scope of this paper, which focuses on the influence of playing technique.
Despite this deliberate restriction, the size of the SOL dataset remains impractically large for collecting human similarity judgments.
Our protocol addresses this problem by means of three complementary approaches: disentanglement of factors, expert pre-screening, and efficient annotation interface.

@@

To construct a dataset for exploring the perception of timbre, we conducted a survey on the perceived similarity of a set of sounds.
The set consists of $78$ recordings of different instrument--PT (IPT) pairs, each played at the same pitch (C4) and nuance \textit{mezzo-forte}. For more information on the dataset considered in this study, see Section \ref{sec:appendix-dataset}.
In the survey, the subject is asked to organize dots on a two-dimensional place in a free sorting task.
Each dot corresponded to a different IPT recording and is moved by the user dragging them with the mouse.
In addition to spatially arranging the dots, the user is also asked to assign one of $20$ colors to each dot, providing a simple information of clustering of the sounds.
While both the location and color of each dot was recorded, only the color is used in our subsequent analysis---the spatial arrangement is only used to facilitate the user's ability to organize the sounds.
More information on the data collection process is available in  Section \ref{sec:subjective}.

@@

\begin{table*}
\caption{Full list of audio stimuli (2/2). In each instrument, a blank space in the rightmost column denotes the ordinary playing technique (\emph{ordinario}).}
\begin{tabular}{llll}
          name & instrument  & mute   &  playing technique \\ \hline \hline
          \texttt{Vn-art-harm-G5-mf-4c} & Violin & & artificial harmonic \\
          \texttt{Vn-legno-batt-C4-mf-4c} & Violin & & col legno battuto \\
          \texttt{Vn-nonvib-C4-mf-4c} & Violin & & nonvibrato \\
          \texttt{Vn-ord-C4-mf-4c} & Violin & & \\
          \texttt{Vn-pizz-bartok-C4-ff-4c} & Violin & & pizzicato Bart\'ok \\
          \texttt{Vn-pizz-lv-C4-mf-4c} & Violin & pizzicato \& \emph{laissez vibrer} \\
          \texttt{Vn-pizz-sec-C4-mf-4c} & Violin & pizzicato secco \\
          \texttt{Vn-pont-C4-mf-4c} & Violin & & sul ponticello \\
          \texttt{Vn+S-ord-C4-mf-4c} & Violin & sordina & \\
          \texttt{Vn+SP-ord-C4-mf-4c} & Violin & piombo & \\ \hline
\end{tabular}
\end{table*}

>>>>

In summary, audio signals from isolated musical notes may vary across three categorical variables (instrument, mute, and technique) and two continuous variables (intensity and pitch).
The set product of all five variables results in a combinatorial explosion of possible timbral sensations. % TODO verify ``set product'' terminology
The Studio On Line dataset (SOL), recorded at Ircam in 2000, offers a dense sampling of this combinatorial explosion, amounting to X audio signals in total. % TODO check year in which SOL was recorded, update with number
We should note that this dense sampling comes at the detriment of other factors of acoustic variability, such as identity of performer, identity of instrument manufacturer, audio acquisition equipment, and room response characteristics; which are all restricted to singletons.
Addressing these factors of variability is besides the scope of this paper, which focuses on the influence of playing technique.
Despite this deliberate restriction, the size of the SOL dataset remains impractically large for collecting human similarity judgments.
Our protocol addresses this problem by means of three complementary approaches: disentanglement of factors, expert pre-screening, and efficient annotation interface.

@@

To construct a dataset for exploring the perception of timbre, we conducted a survey on the perceived similarity of a set of sounds.
The set consists of $78$ recordings of different instrument--PT (IPT) pairs, each played at the same pitch (C4) and nuance \textit{mezzo-forte}. For more information on the dataset considered in this study, see Section \ref{sec:subjective}.
In the survey, the subject is asked to organize dots on a two-dimensional place in a free sorting task.
Each dot corresponded to a different IPT recording and is moved by the user dragging them with the mouse.
In addition to spatially arranging the dots, the user is also asked to assign one of $20$ colors to each dot, providing a simple information of clustering of the sounds.
While both the location and color of each dot was recorded, only the color is used in our subsequent analysis---the spatial arrangement is only used to facilitate the user's ability to organize the sounds.
More information on the data collection process is available in  Section \ref{sec:subjective}.

@@

\begin{table*}
\caption{Full list of audio stimuli (2/2). In each instrument, a blank space in the rightmost column denotes the ordinary playing technique (\emph{ordinario}).}
\label{tab:list2}
\begin{tabular}{llll}
          name & instrument  & mute   &  playing technique \\ \hline \hline
          \texttt{Vn-art-harm-G5-mf-4c} & Violin & & artificial harmonic \\
          \texttt{Vn-legno-batt-C4-mf-4c} & Violin & & col legno battuto \\
          \texttt{Vn-nonvib-C4-mf-4c} & Violin & & nonvibrato \\
          \texttt{Vn-ord-C4-mf-4c} & Violin & & \\
          \texttt{Vn-pizz-bartok-C4-ff-4c} & Violin & & pizzicato Bart\'ok \\
          \texttt{Vn-pizz-lv-C4-mf-4c} & Violin & pizzicato \& \emph{laissez vibrer} \\
          \texttt{Vn-pizz-sec-C4-mf-4c} & Violin & pizzicato secco \\
          \texttt{Vn-pont-C4-mf-4c} & Violin & & sul ponticello \\
          \texttt{Vn+S-ord-C4-mf-4c} & Violin & sordina & \\
          \texttt{Vn+SP-ord-C4-mf-4c} & Violin & piombo & \\ \hline
\end{tabular}
\end{table*}
